Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Just a correction
Starting last lines

We found that the JavaScript ecosystem is mature enough for developing non-trivial 
code, but developers have to consider the differences between VMs implementations
and use certain libraries in order to increase the precision and repeatability of
algorithms. As for the performance of JavaScript against Java and Matlab
we conducted first the basic test described earlier in a 3.7 GHz Quad-Core
Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2,
Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in
Figure \ref{fig:f15_times}. In our test the Matlab implementation had the best
performance with an average of 935ms, followed by Java with
991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a 
single Web Worker and 1,270ms using Node.js. 
The results also show that there is not much difference between the Node.js
implementation an that of a
browser. There is also not much difference between running the code in the main thread or in Web
Workers. When running the experiment in Firefox a dialog alerted that the
script was taking to much time, showing that this
kind of processing is not intended for the main thread. Also there is not much overhead
when running two
Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
language for developing complex mathematical functions for running in the browser. 
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{f15_times.png}
\caption{ Runtime of 10,000 function evaluations (FEs) for $F_{15}$} 
\label{fig:f15_times}
\end{figure}

\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was espected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal aproximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosmes kept in the pool. 
Previously, every PUT request which included a new chromose increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exahust the available memory. In this experiment the size
of a chromosme was 1000 double precision numbers, considereing this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosme was pulled out of the pool before inserting a
new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
and \tt{ classic-ea-float} modules where used. The parameters for the ea algorithm 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosme with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in throu posts in social networks.

@@

@@

In the second implementation, {\sf NodIO$^2$}, Web Workers
were used so that several clients per browser could be executed and
they could run in the background, so that the the number of
generations per user could be improved. The number of generations per
user increased to the point that baseline performance was improved in
most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment at the same time, so that our objective was
achieved. In general, one of the first conclusions for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved with {\sf NodIO$^2$}.

@@

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar way, one
obtained up to 5 times more total cycles  that the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to explore in the future.

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for more computationally demanding
tasks. Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results, but it would be interesting to
measure exactly what this influence has been. Adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.

Finally, the implementation needs some refinement in terms of
programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
to create a generator in which the user just has to create a fitness
function, with the rest of the framework wrapped around
automatically. All this, as the data used for this paper and the paper
itself, has been published with a free license in GitHub at
\url{https://github.com/JJ/modeling-volunteer-computing}.

>>>>

We found that the JavaScript ecosystem is mature enough for developing non-trivial 
code, but developers have to consider the differences between VMs implementations
and use certain libraries in order to increase the precision and repeatability of
algorithms. As for the performance of JavaScript against Java and Matlab
we conducted first the basic test described earlier in a 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2. The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
In our test the Matlab implementation had the best performance with an average of 935ms, followed by Java with
991ms, JavaScript took 32\% more time than Java with 1240ms in Chrome using a single Web Worker and 1,270ms using Node.js. 
The results also show that there is not much difference between the Node.js implementation an that of a
browser. There is also not much difference between running the code in the main thread or in Web Workers. When running the experiment in Firefox a dialog alerted that the script was taking to much time, showing that this
kind of processing is not intended for the main thread. Also there is not much overhead when running two
Web Workers in parallel as they took 1279ms each. This experiment shows that JavaScript is viable 
language for developing complex mathematical functions for running in
the browser. 
%
\begin{figure}[h!tb]
\centering
\includegraphics[width=0.9\linewidth]{f15_times.png}
\caption{ Runtime of 10,000 function evaluations (FEs) for $F_{15}$} 
\label{fig:f15_times}
\end{figure}

\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }

@@

The results for this new experiment have been extremely interesting
and quite different from the previous one.

@@

In the second implementation, {\sf NodIO$^2$}, Web Workers
were used so that several clients per browser could be executed and
they could run in the background, so that the the number of
generations per user could be improved. The number of generations per
user increased to the point that baseline performance was improved in
most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment at the same time, so that our objective was
achieved. In general, one of the first conclusion for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved with {\sf NodIO$^2$}.

@@

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar, way, one
obtained up to 5 times more total cycles  that the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to explore in the future.

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for more computationally demanding
tasks. Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the result, but it would be interesting to
measure exactly what this influence has been. Adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.

Finally, the implementation needs some refinement in terms of
programming and also ease of use. Tools such as Yeoman might be used
to create a generator in which the user just has to create a fitness
function, with the rest of the framework wrapped around
automatically. All this, as the data used for this paper and the paper
itself, has been published with a free license in GitHub at
\url{https://github.com/JJ/modeling-volunteer-computing}.
